TimeVaryingDetection_mrmr

### 2024-11-21 code/R/54188_ClassicCustomComparison_mrmr.Rmd
update introduction

@@

1.  N - total abundance, estimated for each primary period. Included in the standard mrmr model.

2.  B - number of adults recruited into the population, estimated for each primary period. Included in the standard mrmr model.

@@

### Other Blocks...

The custom mrmr model calls upon a Stan file that is modified from the classic mrmr Stan file. Compared to the classic, the modified Stan file...

1.  allows for independent, session-varying estimates of frog detection probabilities

@@

The loo( ) function generates an expected log predictive density (elpd) value for each model. The loo_compare( ) function ranks models based on their elpd values, with the best fit listed first, and calculates the difference in those values. A difference of 2-3x the standard error is generally considered significant in a LOO comparison.

>>>>

The R package mrmr serves as an interface to prepare data and run population models. It relies on a site-specific multi-state Bayesian open-population Jolly-Seber CMR model (hereafter referred to as the multi-state model) for the mathematical and computational part of model fitting.The multi-state model is written in the Stan programming language, which is designed for probabilistic modeling and Bayesian inference. The 'changes to mrmr' outlined in this report are primarily changes to the multi-state model, via modifications to the Stan code. The Stan code is not included in this report. Instead, changes are described using the English language. Snippets of essential Stan code are provided, when necessary, in *italics*.

The multi-state model is composed of seven blocks:

1.  Functions - handle repeated calculations. Key functions are forward_prob( ) and partial_sum_lpmf( ).

2.  Data - specifies input data.

3.  Transformed Data

4.  Parameters - declared the parameters the model will estimate.

5.  Transformed parameters

6.  Model - defines the statistical model and specifies the priors.

7.  Generated quantities - calculates quantities after the model is fit, used for interpretation and diagnostics.

Changes made to each block are described below.

@@

1.  N - total abundance, estimated for each primary period.

2.  B - number of adults recruited into the population, estimated for each primary period.

@@

### Parameters Block

The changes made to this block are essential to achieving independent, time-varying estimates of detection probability.

**beta_detect** a vector of parameters that represent the logit-scale detection probability for survey sessions. The classic mrmr model defines beta_detect as a vector of size m_detect. m_detect refers to the number of detection covariates. The custom model defines beta_detect as a vector of size Jtot. Jtot refers to the total number of survey sessions.

Simply, the original line *vector[m_detect] beta_detect*; is changed to *vector[Jtot] beta_detect;* where each entry corresponds to a specific survey session.

Meaning is assigned to beta_detect in the functions block via transformed parameter logit_detect, within the forward_prob( ) function. No changes were made to this function.

### Transformed Parameters Block

*logit_detect = X_detect \* beta_detect;* —\> *logit_detect = beta_detect;*

In the classic model, logit_detect is derived from beta_detect using the design matrix x_detect. This ties detection probabilities to shared covariates. In the custom model, logit_detect is assigned directly to beta_detect, making each session's detection probability independent. \*

-   When I want to introduce covariates into my analysis, I will need to revisit this section.

@@

The loo( ) function generates an expected log predictive density (elpd) value for each model. The loo_compare( ) function ranks models based on their elpd values, with the best fit listed first, and calculates the difference in those values. A difference of more than 2x the standard error is generally considered significant in a LOO comparison.
